---
title: "Chapter 2 Homework"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r, echo = FALSE}
library(dplyr)
library(rethinking)
library(ggplot2)
```


### 2M1.

**Q:**

Recall the globe tossing model from the chapter. Compute and plot the grid approximate posterior distribution for each of the following sets of observations. In each case, assume a uniform prior for p.

1. [W, W, W]
2. [W, W, W, L]
3. [L, W, W, L, W, W, W]

**Solution:**

W ~ Binomial(N, p)  
p ~ Uniform(0, 1)

Note: *Likelihood* means a distribution function (in this case the binomial distribution) assigned to an observed variable (in this case the W's and L's).

```{r}
# p is our conjectured value, and we construct a grid by distributing it uniformly between 0 and 1.
p_grid <- seq(0, 1, by = 0.01)

# Priors. This encodes the prior plausibility of each value of p.
priors <- rep(1, length(p_grid))

# x and size encode the observed data.
out <- tibble(x = c(3, 3, 5),
              size = c(3, 4, 7)) %>% 
  mutate(likelihoods = purrr::map2(x, size, ~ dbinom(x = .x, size = .y, prob = p_grid)),
         prior = list(priors),
         unstd_posterior = purrr::map2(likelihoods, prior, ~ .x * .y),
         std_posterior = purrr::map(unstd_posterior, ~ .x / sum(.x)),
         p = list(p_grid))

out %>% 
  tidyr::unnest(cols = c(std_posterior, p)) %>% 
  ggplot(aes(p, std_posterior)) +
  geom_line() +
  geom_point() +
  facet_wrap(~ size, ncol = 1)
```


### 2M2.

**Q:**

Now assume a prior for p that is equal to zero when p < 0.5 and is a positive constant when p ≥ 0.5. Again compute and plot the grid approximate posterior distribution for each of the sets of observations in the problem just above.

```{r}
p_grid <- seq(0, 1, by = 0.01)
priors <- ifelse(p_grid < 0.5, 0, 1)

out <- tibble(x = c(3, 3, 5),
              size = c(3, 4, 7)) %>% 
  mutate(likelihoods = purrr::map2(x, size, ~ dbinom(x = .x, size = .y, prob = p_grid)),
         prior = list(priors),
         unstd_posterior = purrr::map2(likelihoods, prior, ~ .x * .y),
         std_posterior = purrr::map(unstd_posterior, ~ .x / sum(.x)),
         p = list(p_grid))

out %>% 
  tidyr::unnest(cols = c(std_posterior, p)) %>% 
  ggplot(aes(p, std_posterior)) +
  geom_line() +
  geom_point() +
  facet_wrap(~ size, ncol = 1)
```


### 2M3.

**Q:**

```{r}

```














